---
title: "HW6 - FunctionsAndLoops"
author: "Malte Højmark-Bertelsen"
date: "22/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#  Explore the recent global developments with R
Today, you will load a filtered gapminder dataset -  with a subset of data on global development from 1952 - 2007 in increments of 5 years - to capture the period between the Second World War and the Global Financial Crisis. 

## Get the necessary packages
First, start with installing the relevant packages 'tidyverse', 'gganimate', and 'gapminder'.

```{r libraries, echo = FALSE}
library(pacman)
p_load(tidyverse, gganimate, gapminder)
```

# 1. Define a defensive function that calculates the Gross Domestic Product of a nation from the data available in the gapminder dataset. Using that function, calculate the GDP of Denmark in the following years: 1967, 1977, 1987, 1997, 2007, and 2017.

## Creating a defensive function
```{r}
calcGDP <- function(dat, year=NULL, country=NULL) {
  if(!is.null(year)) {
    dat <- dat[dat$year %in% year, ]
  }
  if (!is.null(country)) {
    dat <- dat[dat$country %in% country,]
  }
  gdp <- dat$pop * dat$gdpPercap

  new <- cbind(dat, gdp=gdp)
  return(new)
}

```
Here in this function it is optional whether to include a year and a country. If including these variables, the function will subset the data containing only the specified year and country. Afterwards the GDP will be created and concatenate the dataframe with the gdp for the given year and country.

# Printing years for Denmark
```{r}
df = as.data.frame(gapminder)
years = c(1967, 1977, 1987, 1997, 2007, 2017)
for (year in years){
  row = calcGDP(dat = df, year = year, country = "Denmark")
  print(row)
}
```
Here i create a dataframe-holder "df" and a year-holder "years" and then i loop through each year, calculate the gpd for Denmark and print the specific row. 

# 2. Write a script that loops over each country in the gapminder dataset, tests whether the country starts with a ‘B’ , and print out whether the life expectancy is smaller than 50, between 50 and 70, or greater than 70.

##Loop for printing life expectancy
```{r}
for (country in unique(gapminder$country)){
  if (substring(country, 1, 1) == "B"){
    temp_df = df[df$country %in% country,]
    for (year in unique(temp_df$year)){
      temp_df_year = temp_df[temp_df$year %in% year,]
      if (temp_df_year$lifeExp < 50){
        cat("For the year of", year, "the country", country, "had a life expectancy below 50. \n")
      }
      else if (temp_df_year$lifeExp > 50 && temp_df_year$lifeExp < 70){
        cat("For the year of", year, "the country", country, "had a life expectancy between 50 and 70. \n")
      }
      else{
        cat("For the year of", year, "the country", country, "had a life expectancy above 70. \n")
      }
    }
  }
}

```
Here i check to see if each unique country in the dataframe starts with a "B". Then i subset the given country from the data. Afterwards i subset it by the unique years the country has represented. Afterwards i create print statements with the conditional rules provided in the assignment description.