---
title: "Meat consumption data scraping"
author: "Malte Højmark-Bertelsen"
date: "20 July 2020, updated `r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Solution
First, install a handful of classic R packages and load their libraries:

- `rvest` for web-scraping
- `dplyr` for data-wrangling
- `tidyr` for data transformation
- `stringr` for string manipulation
- `janitor` for clean headers that your OCD will love you for


```{r libraries}
library(rvest)
library(dplyr)
library(tidyr)
library(stringr)
library(janitor)
```


## Scrape the data

We start by defining an URL and reading the HTML from the URL.
```{r}
url <- "https://en.wikipedia.org/wiki/List_of_countries_by_meat_consumption"
# scrape the website
url_html <- read_html(url)
```

Afterwards, we use load the data from a specific HTML-node into a table, and create a tibble.

```{r, option3}
whole_table <- url_html %>% 
 html_node("table") %>%
 html_table(fill = TRUE)  #str(whole_table) turns out to be a list

head(whole_table)
whole_table %>% head() #The data scraped from Wiki

meat <- as_tibble(whole_table)
meat %>% head() #The data scraped as a tibble

```
Defining a function that does all the work above for us
```{r}
scrape_meat_consumption_wiki <- function(website){
	url <- read_html(website)
	list_table <- url %>% 
 			html_node("table") %>%
 			html_table(fill = TRUE)  # result is a list
  tibble <- as_tibble(list_table)
  
}

url <- "https://en.wikipedia.org/wiki/List_of_countries_by_meat_consumption"
meat <- scrape_meat_consumption_wiki(url)
meat %>% head()
```



# Cleaning
The column names has remnants from Wikipedia source links, which is why i will rename these.
```{r}
meat <- meat %>% rename('kg_per_person_2002' = 'kg/person (2002)[9][note 1]')
meat <- meat %>% rename('kg_per_person_2009' = 'kg/person (2009)[10]')
meat <- meat %>% rename('kg_per_person_2017' = 'kg/person (2017)[11]')
meat %>% head()


```
And there you have a dataframe with countries and their corresponding meat consumption for the years 2002, 2009 and 2017 :-)
